---
title: "Introduction to WDC Connector"
author: "Angus Moir (angus.moir@bankofengland.co.uk)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package provides a set of tools that allows R users to dynamically serve up tables and connect them to Tableau using Tableau's [Web Data Connector](https://tableau.github.io/webdataconnector/).

To understand how it works we are going to serve up two classic R tables, `mtcars` and `iris` and connect to them in Tableau.

## Quick start

The quickest way to see the `wdcconnector` package in action is with an example:

Tableau doesn't accept columns names containing full stops so we need to rename our `iris` columns.
```{r}
iris <- dplyr::rename(iris,
                      Sepal_Length = Sepal.Length,
                      Sepal_Width = Sepal.Width,
                      Petal_Length=Petal.Length,
                      Petal_Width=Petal.Width)

```

Let's generate some file paths that will be used so server our various files.

```{r}
js_wdc_server_path <- "wdcjs.js"
js_wdc_dir_path <- file.path(here::here("examples"), "simple", js_wdc_server_path)

html_path <- here::here("examples/simple/wdchtml.html")
plumbed_file_path <- here::here("examples/simple/plumbed_file.R")
```

Create the javascript for our iris Web Data Connector
```{r}
wdcjs <- generateWDCJS(iris)
wdcjs
```

And save that to disk
```{r}
writeLines(wdcjs, js_wdc_dir_path)
```

Create the WDC HMTL file
```{r}
wdchtml <- generateWDCHTML(js_wdc_server_path)
wdchtml

```

And write it to disk
```{r}
writeLines(wdchtml, html_path)
```

Generate the file to be plumbed.
```{r}
writeLines(generatePlumbedFile(iris), plumbed_file_path)
```

Finally we need to serve the various file
```{r}
r <- plumber::plumb(plumbed_file_path)

web_connector_html <- plumber::PlumberStatic$new("examples/simple")

r$mount("/wdc", web_connector_html)

r$run(port=8000)

```


Open up your version of Tableau and choose the Web Data Connector option.

Copy and paste the following URL into the address bar:

http://127.0.0.1:8000/wdchtml.html

Click "Get R Data" when it becomes available.

Update your data.

## Overview of the Tableau Web Data Connector

In order for Tableau to connect to a table in R dynamically, you need to build a [Web Data Connector](https://tableau.github.io/webdataconnector/) for that table. 

This web data connector is a javascript object that contains 2 key user defined methods:

- `getSchema` is going to define the column names, types and table name of your table.
- `getData` is going to determine how Tableau should import table from your data served as a JSON into the table.

To dynamically generate the `getSchema` method the `wdcconnector` package has a function:
```{r}
generateWDCGetSchemaJS(mtcars)
```

To dynamically generate the `getData` method the `wdcconnector` package has a function:
```{r}
generateWDCDataJS(mtcars)
```

Finally you need to register the web data connector as an object you have created and link it to the HTML interface.

This can be generated using the function:
```{r}
generateWDCConnectorJS()
```

